\documentclass[10pt]{article}
\usepackage{amsmath,amssymb,bm,fullpage,relsize}
\usepackage{multicol}
\usepackage{geometry}
\geometry{
	top = 1.5cm,
	left = 1.75cm,
	right = 1.75cm
}

\newcommand{\x}{\bm{x}}
\renewcommand{\a}{\bm{a}}
\renewcommand{\b}{\bm{b}}
\renewcommand{\u}{\bm{u}}
\renewcommand{\v}{\bm{v}}
\newcommand{\R}{{\mathbb R}}
\newcommand{\zero}{\bm{0}}

\begin{document}

\begin{center}
{\Large \textbf{Chase Joyner}}
\end{center}
O-110 Martin Hall, Box 340975 \hfill chasej@g.clemson.edu \\
Clemson, SC 29634 \hfill (704) 830-4730 \\
\rule{18.08cm}{0.20mm} \\ \\
\noindent
\textbf{Education}
\begin{itemize}
\item[-] Ph.D. in Mathematical Sciences (Statistics) (4.0/4.0) \hfill May 2016 -- Present \\
Clemson University, Clemson, SC
\item[-] M.S. in Mathematical Sciences (Statistics) (4.0/4.0) \hfill Aug 2014 -- May 2016 \\
Clemson University, Clemson, SC
\item[-] B.S. in Mathematical Sciences (Statistics) (3.89/4.0) \hfill Aug 2010 -- 2014 \\
Clemson University, Clemson, SC
\end{itemize}

\noindent
\textbf{Research Experience: Clemson University}
\begin{itemize}
\item[-] Research Assistant sponsored by National Institute of Health \hfill Jan 2017 -- Present

\vspace{-0.25cm}

\begin{itemize}
\item[-] Developed a Bayesian mixed effects model with variable selection to analyze \\ data 
from any group testing algorithm while accounting for imperfect testing.

\vspace{-0.05cm}

\item[-] Analyzed a chlamydia data set out of Iowa's state hygienic laboratory.

\vspace{-0.05cm}

\item[-] Techniques used: Generalized linear models, spike and slab priors.
\end{itemize}
\item[-] Research Assistant sponsored by Biorealm, Principal Investigator \hfill Jan 2016 -- Dec 2016
\begin{itemize}

\vspace{-0.25cm}

\item[-] Analyzed rice data provided from fields in Indonesia to develop a mixed effects \\ model accounting for complex genetic similarity.

\vspace{-0.05cm}

\item[-] Modeled rice production and resistance to climate change in Indonesia.

\vspace{-0.05cm}

\item[-] Simulated data in R using Clemson's cluster to validate the model.

\vspace{-0.05cm}

\item[-] Techniques used: Expectation-Maximization algorithm, generalized linear models, \\ and mixed effects models.
\end{itemize}
\item[-] Master's Thesis \hfill Aug 2014 -- May 2016
\begin{itemize}

\vspace{-0.25cm}

\item[-] Developed univariate and multivariate Bayesian models to estimate the optimal \\ biomarker density threshold in pooled testing of individuals for various diseases.

\vspace{-0.05cm}

\item[-] Implemented algorithms in R to estimate the parameters of these Bayesian models.

\vspace{-0.05cm}

\item[-] Techniques used: Gibbs sampling, Metropolis-Hastings, and Bayesian iteratively \\ reweighted least squares.
\end{itemize}
\item[-] Undergraduate Thesis \hfill Aug 2013 -- Aug 2014
\begin{itemize}

\vspace{-0.25cm}

\item[-] Analyzed Bayesian techniques and Markov chain Monte Carlo methods for inference.

\vspace{-0.05cm}

\item[-] Documented the implementation of these methods and ran simulations.
\end{itemize}
\end{itemize}

\noindent
\textbf{Research Presentations}
\begin{itemize}
\item[-] A mixed effects model for group testing data with variable selection. \\
\emph{ENAR Spring Meeting, Atlanta, GA} (Mar 2018).
\item[-] Assessing the relationship between SNPs and yield in various rice varieties. \\
\emph{Jakarta, Indonesia} (Nov 2016).
\item[-] Bayesian approach of biomarker density estimation using pooled data. \\
\emph{Clemson University} (Feb 2016).
\end{itemize}

\newpage
\noindent
\textbf{Publications}
\begin{itemize}
\item[-] Joyner, C., McMahan, C., Tebbs, J., and Bilder, C. (2018+).  From mixed effects modeling to spike and slab variable selection: A Bayesian regression model for group testing data. In preparation.
\item[-] McMahan, C., Baurley, J., Bridges, W., Joyner, C., Fitra Kacamarga, M., Lund, R., Pardamean, C., and Pardamean, B. (2017). A Bayesian hierarchical model for identifying significant polygenic effects while controlling for confounding and repeated measures. \textit{Statistical Applications in Genetics and Molecular Biology}. 16, 407-419.
\end{itemize}

\noindent
\textbf{Teaching Experience}
\begin{itemize}
\item[-] BINUS University, Department of Mathematical Statistics, Jakarta, Indonesia \hfill May 2018
\begin{itemize}

\vspace{-0.25cm}

\item[-] Introduction to statistics and R
\end{itemize}
\item[-] Clemson University, Departmental of Mathematical Sciences, Clemson, SC
\begin{itemize}

\vspace{-0.25cm}

\item[-] MATH 1070: Differential and Integral Calculus \hfill Spring 2017

\vspace{-0.05cm}

\item[-] MATH 1040: Precalculus and Introductory Differential Calculus \hfill Fall 2016

\vspace{-0.05cm}

\item[-] MATH 1020: Introduction to Mathematical Analysis \hfill Fall 2015
\end{itemize}
\end{itemize}

\noindent
\textbf{Professional Memberships}
\begin{itemize}
\item[-] American Mathematical Society (AMS)
\end{itemize}

\noindent
\textbf{References}
\begin{itemize}
\item[-] Christopher McMahan, Clemson University. Contact: \emph{mcmaha2@g.clemson.edu}.
\item[-] Andrew Brown, Clemson University. Contact: \emph{ab7@g.clemson.edu}.
\end{itemize}

\end{document}
